---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(readxl)
library(here) # for getting the working directory
library(lme4)
library(tidyverse)
library(nlme)
```

This is where I add the ontogeny identifiers "juvenile", "transition", and "adult"
```{r}
here()

#Morphology
PoacherMorph <- read_excel("PoacherArmorDataCMD.xlsx", sheet = "mashed data(kin_morph)")
  PoacherMorph <- mutate(PoacherMorph, length = length/100)
  PoacherMorph$bins <- cut(PoacherMorph$length, breaks=c(0,0.06,0.105,0.5), labels=c("juvenile","transition","adult"))

#Kinematics
PoacherKinematics <- read_csv("PoacherKinematics.csv")
  PoacherKinematics <- mutate(PoacherKinematics, length = length/1000)
  PoacherKinematics$bins <- cut(PoacherKinematics$length, breaks=c(0,0.06,0.105,0.5),
                                labels=c("juvenile","transition","adult"))
    #Reynolds number stuff
    density <- 1027
    viscosity <- 0.00141
    dWarm <- 1023	
    viscWarm <- 0.00096
  PoacherKinematics <- mutate(PoacherKinematics, Re = ((density*speed*length)/viscosity))
  PoacherKinematics <- mutate(PoacherKinematics, ReWarm = ((dWarm*speed*length)/viscWarm))

#Bending Mechanics
MTSmerge<- read_csv("PoacherMTSMerge.csv")
  MTSmerge <- mutate(MTSmerge, length = length/1000)
  MTSmerge <- mutate(MTSmerge, stress = stress/10^6)
  MTSmerge <- mutate(MTSmerge, stiffness = stiffness*(1/1000))
  MTSmerge$bins <- cut(MTSmerge$length, breaks=c(0,0.06,0.105,0.5), labels=c("juvenile","transition","adult"))

#Kinematics + Mechanics
Mechanics <- merge(PoacherKinematics, MTSmerge,by="indiv")

#Scale Density
ScaleDensity02 <- read_csv('NewNewScaleData2020_02.csv')
  ScaleDensity02$bins <- cut(ScaleDensity02$size, breaks=c(0,60,105,500),
                           labels=c("juvenile","transition","adult"))
  
  
  
ScaleDensityMeans <- aggregate(ScaleDensity02, list(ScaleDensity02$Individual, ScaleDensity02$size), mean)
ScaleDensityMeans$bins <- cut(ScaleDensityMeans$size, breaks=c(0,60,105,500),
                           labels=c("juvenile","transition","adult"))

ScaleDen02 <- ScaleDensity02[ScaleDensity02$Individual != "Avulsa_XL",]
MeanScaleDen02 <- ScaleDensityMeans[ScaleDensityMeans$Group.1 != "Avulsa_XL",]

```

Morphology stuff
Calculating average height, width, length, area, and volume of the scales
```{r}
PoacherMorph <- mutate(PoacherMorph, R1a = R1w*R1l)
PoacherMorph <- mutate(PoacherMorph, R2a = R2w*R2l)
PoacherMorph <- mutate(PoacherMorph, R3a = R3w*R3l)
PoacherMorph <- mutate(PoacherMorph, R4a = R4w*R4l)
PoacherMorph <- mutate(PoacherMorph, R5a = R5w*R5l)
PoacherMorph <- mutate(PoacherMorph, R6a = R6w*R6l)
PoacherMorph <- mutate(PoacherMorph, R7a = R7w*R7l)
PoacherMorph <- mutate(PoacherMorph, R8a = R8w*R8l)

PoacherMorph <- mutate(PoacherMorph, R1v = R1a*R1h)
PoacherMorph <- mutate(PoacherMorph, R2v = R2a*R2h)
PoacherMorph <- mutate(PoacherMorph, R3v = R3a*R3h)
PoacherMorph <- mutate(PoacherMorph, R4v = R4a*R4h)
PoacherMorph <- mutate(PoacherMorph, R5v = R5a*R5h)
PoacherMorph <- mutate(PoacherMorph, R6v = R6a*R6h)
PoacherMorph <- mutate(PoacherMorph, R7v = R7a*R7h)
PoacherMorph <- mutate(PoacherMorph, R8v = R8a*R8h)

PoacherMorph <- mutate(PoacherMorph, MeanL = rowMeans(cbind(R1l, R2l, R3l, R4l, R5l, R6l, R7l, R8l), na.rm = TRUE))
PoacherMorph <- mutate(PoacherMorph, MeanW = rowMeans(cbind(R1w, R2w, R3w, R4w, R5w, R6w, R7w, R8w), na.rm = TRUE))
PoacherMorph <- mutate(PoacherMorph, MeanH = rowMeans(cbind(R1h, R2h, R3h, R4h, R5h, R6h, R7h, R8h), na.rm = TRUE))
PoacherMorph <- mutate(PoacherMorph, MeanA = rowMeans(cbind(R1a, R2a, R3a, R4a, R5a, R6a, R7a, R8a), na.rm = TRUE))
PoacherMorph <- mutate(PoacherMorph, MeanV = rowMeans(cbind(R1v, R2v, R3v, R4v, R5v, R6v, R7v, R8v), na.rm = TRUE))

PoacherMorph <- mutate(PoacherMorph, sdL = apply(cbind(R1l, R2l, R3l, R4l, R5l, R6l, R7l, R8l), 1, sd, na.rm = TRUE))
PoacherMorph <- mutate(PoacherMorph, sdW = apply(cbind(R1w, R2w, R3w, R4w, R5w, R6w, R7w, R8w), 1, sd, na.rm = TRUE))
PoacherMorph <- mutate(PoacherMorph, sdH = apply(cbind(R1h, R2h, R3h, R4h, R5h, R6h, R7h, R8h), 1, sd, na.rm = TRUE))
PoacherMorph <- mutate(PoacherMorph, sdA = apply(cbind(R1a, R2a, R3a, R4a, R5a, R6a, R7a, R8a), 1, sd, na.rm = TRUE))
PoacherMorph <- mutate(PoacherMorph, sdV = apply(cbind(R1v, R2v, R3v, R4v, R5v, R6v, R7v, R8v), 1, sd, na.rm = TRUE))

PoacherMorph <- mutate(PoacherMorph, MeanLBL = MeanL/length)
PoacherMorph <- mutate(PoacherMorph, MeanWBL = MeanW/length)
PoacherMorph <- mutate(PoacherMorph, MeanHBL = MeanH/length)
PoacherMorph <- mutate(PoacherMorph, MeanABL = MeanA/length^2)
PoacherMorph <- mutate(PoacherMorph, MeanVBL = MeanV/length^3)
PoacherMorph <- mutate(PoacherMorph, MeanAR = MeanW/MeanL)
```

Scale Density Stuff

scaleDenModel02 Results
Response: bmd
          Df  Sum Sq Mean Sq F value Pr(>F)
bins       2  110.76  55.381  2.2243 0.1183
scale      5    7.67   1.535  0.0617 0.9973
Residuals 52 1294.72  24.898   

```{r}
# scaleDenModel1 <- lm(bmd ~ bins + scale, random = ~1 | Individual, data = ScaleDensity02)
# anova(scaleDenModel1)
# 
scaleDenModel2 <- lm(bmd ~ bins + scale, random = ~1 | Individual, data = ScaleDen02)
summary(scaleDenModel2)

scaleDenModel3 <- lm(bmd ~ bins, random = ~1 | Group.1, data = ScaleDensityMeans)
summary(scaleDenModel3)

scaleDenModel4 <- lm(bmd ~ bins, random = ~1 | Group.1, data = MeanScaleDen02)
anova(scaleDenModel4)

# plot(scaleDenModel)
# plot(scaleDenModel2)

ggplot(ScaleDensityMeans, aes(x = bins, y = bmd, color = bins))+
  geom_boxplot()+
  theme_classic()
# 
# ggplot(ScaleDensity, aes(x = log10(size), y = bmd, color = as.factor(Individual)))+
#   geom_point()+
#   theme_classic()

ggplot(MeanScaleDen02, aes(x = bins, y = bmd, color = bins))+
  geom_boxplot()+
  theme_classic()
#ggsave("ScaleDensity.eps", width = 15, height = 10, units = "cm")

```

Stiffness Interpolation
```{r}
mechSplit <- split(Mechanics, Mechanics$bins.x)
Juveniles <- mechSplit$juvenile
Transition <- mechSplit$transition
Adult <- mechSplit$adult

JuvStiffFun <- approxfun(Juveniles$bendingPts, Juveniles$stiffness, rule = 2, method = "constant")
TraStiffFun <- approxfun(Transition$bendingPts, Transition$stiffness, rule = 2, method = "constant")
AduStiffFun <- approxfun(Adult$bendingPts, Adult$stiffness, rule = 2, method = "constant")

morphSplit <- split(PoacherMorph, PoacherMorph$bins)
JuvMorph <- morphSplit$juvenile
TraMorph <- morphSplit$transition
AduMorph <- morphSplit$adult

JuvMorph <- mutate(JuvMorph, stiffInterp = JuvStiffFun(JuvMorph$position))
TraMorph <- mutate(TraMorph, stiffInterp = TraStiffFun(TraMorph$position))
AduMorph <- mutate(AduMorph, stiffInterp = AduStiffFun(AduMorph$position))

newPoacherMorph <- rbind(JuvMorph, TraMorph, AduMorph)

# my_cols <- c("deeppink2", "chartreuse2", "deepskyblue1")
# pairs(newPoacherMorph[,c(2:3,45:46, 48, 55:56)],
#       col = my_cols[as.factor(newPoacherMorph$bins)])

allStiffFun <- approxfun(Mechanics$length.x, Mechanics$stiffness, rule = 2)
PoacherMorph <- mutate(PoacherMorph, stiffInterpAll = allStiffFun(PoacherMorph$length))

```

Stiffness Interpolation Plots
```{r}
ggplot(newPoacherMorph, aes(x = MeanLBL, y = stiffInterp))+
  stat_smooth(aes(x = MeanLBL, y = stiffInterp), method = "lm", se = F)+
  geom_point(aes(color = bins))+  
  theme_classic()

ggplot(newPoacherMorph, aes(x = MeanWBL, y = stiffInterp))+
  stat_smooth(aes(x = MeanWBL, y = stiffInterp), method = "lm", se = F)+
  geom_point(aes(color = bins))+  
  theme_classic()

ggplot(newPoacherMorph, aes(x = MeanHBL, y = stiffInterp))+
  stat_smooth(aes(x = MeanHBL, y = stiffInterp), method = "lm", se = F)+
  geom_point(aes(color = bins))+  
  theme_classic()
```



Looking at individual variables
- Mechanics supports two groups (Juvenile and Adult)
- Morphology supports three groups
- Kinematics --> Juveniles beat their tails faster and generally swim faster (in BL/second)
- They are less stiff then adults, but the same as "transitional" individuals
- They tend to have shorder and wider plates at their tails then adults (which could increase flexiblilty)


Kinematics Figures
```{r}
 #fits <- lmList(bendingPts ~ stiffness | indiv, data=MTSmerge)
 #fits <- lmList(bendingPts ~ stress | indiv, data=MTSmerge)
 
ggplot(PoacherKinematics)+
  stat_smooth(aes(x = length, y = amplitude/length), method = "lm", se = F)+
  geom_point(aes(x = length, y = amplitude/length, color = bins))+  
  geom_smooth(aes(x = length, y = amplitude/length, color = bins), method=lm, se=FALSE)+
  theme_minimal()
#ggsave("SLvAMP.eps", width = 10, height = 5, units = "cm")

 ggplot(PoacherKinematics)+
  stat_smooth(aes(x = length, y = speed/length), method = "lm", se = F)+
  geom_point(aes(x = length, y = speed/length, color = bins))+  
  geom_smooth(aes(x = length, y = speed/length, color = bins), method=lm, se=FALSE)+
  theme_minimal()
#ggsave("SLvSPEED.eps", width = 10, height = 5, units = "cm")
 
ggplot(PoacherKinematics)+
  stat_smooth(aes(x = length, y = stride/length), method = "lm", se = F)+
  geom_point(aes(x = length, y = stride/length, color = bins))+ 
  geom_smooth(aes(x = length, y = stride/length, color = bins), method=lm, se=FALSE)+
  theme_minimal()
#ggsave("SLvSTRIDE.eps", width = 10, height = 5, units = "cm")
  
ggplot(PoacherKinematics)+
  stat_smooth(aes(x = length, y = frequency), method = "lm", se = F)+
  geom_point(aes(x = length, y = frequency, color = bins))+  
  geom_smooth(aes(x = length, y = frequency, color = bins), method=lm, se=FALSE)+
  theme_minimal()
#ggsave("SLvFREQ.eps", width = 10, height = 5, units = "cm")
```

Morphology Along the Body Figures
```{r}
ggplot(PoacherMorph)+
  geom_errorbar(aes(x = length, y = MeanL, group = bins, color = bins, ymin = MeanL-sdL, ymax = MeanL+sdL), width = 0.01)+
  geom_point(aes(x = length, y = MeanL, group = bins, color = bins, size = position))+
  theme_classic()+
  geom_smooth(aes(x = length, y = MeanL), method = "lm", se = FALSE)
ggsave("LvsL.eps", width = 15, height = 10, units = "cm")

ggplot(PoacherMorph)+
  geom_errorbar(aes(x = length, y = MeanW, group = bins, color = bins, ymin = MeanW-sdL, ymax = MeanW+sdL), width = 0.01)+
  geom_point(aes(x = length, y = MeanW, group = bins, color = bins, size = position))+
  theme_classic()+
  geom_smooth(aes(x = length, y = MeanW), method = "lm", se = FALSE)
ggsave("LvsW.eps", width = 15, height = 10, units = "cm")

ggplot(PoacherMorph)+
  geom_errorbar(aes(x = length, y = MeanH, group = bins, color = bins, ymin = MeanH-sdL, ymax = MeanH+sdL), width = 0.01)+
  geom_point(aes(x = length, y = MeanH, group = bins, color = bins, size = position))+
  theme_classic()+
  geom_smooth(aes(x = length, y = MeanH), method = "lm", se = FALSE)
ggsave("LvsH.eps", width = 15, height = 10, units = "cm")

ggplot(PoacherMorph)+
  geom_errorbar(aes(x = length, y = MeanA, group = bins, color = bins, ymin = MeanA-sdL, ymax = MeanA+sdL), width = 0.01)+
  geom_point(aes(x = length, y = MeanA, group = bins, color = bins, size = position))+
  theme_classic()+
  geom_smooth(aes(x = length, y = MeanA), method = "lm", se = FALSE)
ggsave("LvsA.eps", width = 15, height = 10, units = "cm")

ggplot(Mechanics, aes(x = length.x, y = stiffness, group = bins.x, color = bins.x))+
  geom_boxplot()+
  theme_classic()
#ggsave("LvsSTIFF.eps", width = 10, height = 5, units = "cm")

```

Tessa's figures for ICVM
```{r}
ggplot(PoacherMorph, aes(x = length, y = MeanHBL, color = bins))+
  geom_point()+ 
  geom_smooth(method=lm, se=FALSE) + theme(panel.background = element_rect(fill = "black"),
         plot.background = element_rect(fill="black"),
         axis.title.x = element_text(colour = "white"),
         axis.title.y = element_text(colour = "white"),
         legend.text = element_text(colour = "white"),
         legend.background = element_rect(fill = "black"))

ggplot(MTSmerge, aes(x = bendingPts, y = stiffness, color = bins))+
  geom_point()+ 
  geom_smooth(method=lm, se=FALSE) + theme(panel.background = element_rect(fill = "black"),    plot.background = element_rect(fill="black"),
         axis.title.x = element_text(colour = "white"),
         axis.title.y = element_text(colour = "white"),
         legend.text = element_text(colour = "white"),
         legend.background = element_rect(fill = "black"))
```

Linear Stats (Morphology)
TL;DR - relative length and width don't really change with ontogeny but spine height does
```{r}
AreaLM <- lm(MeanABL ~ length, data = PoacherMorph)
summary(AreaLM)

HeightLM <- lm(MeanHBL ~ length, data = PoacherMorph)
summary(HeightLM)

LengthLM <- lm(MeanLBL ~ length, data = PoacherMorph)
summary(LengthLM)

WidthLM <- lm(MeanWBL ~ length, data = PoacherMorph)
summary(WidthLM)
```

Linear Stats (Swimming)
TL;DR - frequency and relative speed change with ontogeny but relative amp and SL do not
```{r}
AmpLM <- lm((amplitude/length) ~ length, data = PoacherKinematics)
summary(AmpLM)

SpeedLM <- lm((speed/length) ~ length, data = PoacherKinematics)
summary(SpeedLM)

StrideLM <- lm((stride/length) ~ length, data = PoacherKinematics)
summary(StrideLM)

FreqLM <- lm(frequency ~ length, data = PoacherKinematics)
summary(FreqLM)
```

Linear Stats (Mechanics)
```{r}
stiffnessLM <- lm(stiffness ~ length, data = MTSmerge)
summary(stiffnessLM)

ggplot(MTSmerge, aes(x = length, y = stiffness, color = bendingPts))+
  geom_point()
```

Scailing Stats (Kinematics)
```{r}
require(lmodel2)
ampRMA <- lmodel2(formula = log(amplitude) ~ log(length), data = PoacherKinematics, nperm = 99)
ampRMA  

strideRMA <- lmodel2(formula = log(stride) ~ log(length), data = PoacherKinematics, nperm = 99)
strideRMA 

speedRMA <- lmodel2(formula = log(speed) ~ log(length), data = PoacherKinematics, nperm = 99)
speedRMA 

freqRMA <- lmodel2(formula = log(frequency) ~ log(length), data = PoacherKinematics, nperm = 99)
freqRMA 
```

Reynolds Number Figure
```{r}
ggplot(PoacherKinematics, aes(length, Re, color = bins))+
  geom_point()+
  geom_hline(yintercept=2300)+
  geom_hline(yintercept=4000)+
  theme_classic()
#ggsave("LvRE.eps", width = 10, height = 5, units = "cm")

# ggplot(PoacherKinematics, aes(length, ReWarm, color = bins))+
#   geom_point()+
#   geom_hline(yintercept=1000)+
#   geom_hline(yintercept=10000)
```

PCA Stuff
```{r}
library(ggfortify)
library(factoextra)
library("MASS")
library("Momocs")

res1.man <- manova(cbind(MeanABL, MeanWBL, MeanHBL, MeanLBL) ~ bins, data = PoacherMorph)
summary(res1.man)
summary.aov(res1.man)

res2.man <- manova(cbind(amplitude/length, speed/length, stride/length, frequency) ~ bins, data = PoacherKinematics)
summary(res2.man)
summary.aov(res2.man)

res3.man <- manova(cbind(stiffness, stress) ~ bins, data = MTSmerge)
summary(res3.man)
summary.aov(res3.man)

morph.pca <- prcomp(PoacherMorph[,c(55:59)], center = TRUE,scale. = TRUE)
morph.p <- PCA(PoacherMorph[,c(2, 55:59)], center = TRUE, scale. = TRUE)
morph.lda <- lda(PoacherMorph$bins ~ PoacherMorph$MeanABL + PoacherMorph$MeanVBL + PoacherMorph$MeanLBL +
                   PoacherMorph$MeanWBL + PoacherMorph$MeanHBL)
morph.lda.values <- predict(morph.lda)
ldahist(data = morph.lda.values$x[,1], PoacherMorph$bins)

fviz_pca_biplot(morph.pca, label="var", habillage=PoacherMorph$bins,
               addEllipses=TRUE, ellipse.level=0.5,col.var = "yellow")+ 
  labs(x = "PC1", y = "PC2")+
  theme_classic()
ggsave("morphPCA.eps", width = 15, height = 10, units = "cm")

kine.pca <- prcomp(PoacherKinematics[,c(5:8,10)], center = TRUE,scale. = TRUE)
kine.p <- PCA(PoacherKinematics[,c(3, 5:8,10)], center = TRUE, scale. = TRUE)
kine.lda <- lda(PoacherKinematics$bins ~ + PoacherKinematics$speed + PoacherKinematics$frequency +
                  PoacherKinematics$stride + PoacherKinematics$amplitude)
kine.lda.values <- predict(kine.lda)
ldahist(data = kine.lda.values$x[,1], PoacherKinematics$bins)

fviz_pca_biplot(kine.pca, label="var", habillage=PoacherKinematics$bins,
               addEllipses=TRUE, ellipse.level=0.5, col.var = "yellow")+
  labs(x = "PC1", y = "PC2")+
   theme_classic()
ggsave("kinePCA.eps", width = 15, height = 10, units = "cm")
```

Look at all the morphology data
Conclusions (July)
1) Everything increases as poacher length increases, but the slopes are different
2) Position along the body matters (looks pretty linear so we can fit a line to it)
3) Area and volume increase from tail to head (this looks like it's mostly due to width and height)
```{r}
my_cols <- c("deeppink2", "chartreuse2", "deepskyblue1") 
pairs(PoacherMorph[,c(2,44:48)],
      col = my_cols[as.factor(PoacherMorph$position)])

pairs(PoacherMorph[,c(2,49:51)],
      col = my_cols[as.factor(PoacherMorph$position)])
```

Things to do:
- Re run stats without the groups (provide evidence for groups)
- How does naure build a flexible armored tube
  - how are plates arranged along the body
  - Do the tiles vary along the body
  - Do the number of tiles change with ontogeny
  - How do changes in tiles change the stiffness of the tube
